---
title: "Reproducible Research Project 1"
output: html_document
date: "2023-02-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
#read the csv file
data<-read.csv("activity.csv")
#date is stored as char, change to date type
data$date<-as.Date(data$date)
#calculate total steps each day
totalperday<-aggregate(data$steps, list(data$date), FUN=sum)
```

## Histogram of the total number of steps taken each day
```{r}
#plot histogram
hist(totalperday$x, main="Total Steps Taken Each Day", xlab="Total Steps in a Day", col="red", border="black")

```

## Mean 
```{r}
 mean(totalperday$x, na.rm=TRUE)
```
## Median
```{r}
median(totalperday$x, na.rm=TRUE)
```

## Time Series Plot of the 5-Minute interval
```{r}
#calculate average steps across all days
avg_steps_interval<-aggregate(steps~interval, data, FUN=mean)
#plot the graph
plot(avg_steps_interval$interval, avg_steps_interval$steps, type="l", main="Average Number of Steps By 5 Minute Interval",xlab="5-minute Interval",ylab="Average Number of Steps")
```

## 5-Minute interval with the maximum number of steps
```{r}
max_index<-which.max(avg_steps_interval$steps)
avg_steps_interval[max_index,]
```
# Imputing missing values
## Total number of missing values in the dataset is
```{r}
sum(is.na(data$steps))
```
## Filling in NAs with the mean of that 5 minute interval value
```{r}
#use for loop to overwrite NAs with the mean of that 5 minute interval.
for (i in 1:nrow(data)){
  if(is.na(data$steps[i])){
    data$steps[i]<-avg_steps_interval$steps[which(avg_steps_interval$interval==data$interval[i])]
   }
}
```
```{r}
#aggregate the steps taken each day with the new imputed data
totalperday_impute<-aggregate (steps~date, data, sum)

#plot the histogram with Imputed data, after filling in missing valuse (NAs)
hist(totalperday_impute$steps, main="Total Steps Taken Each Day (IMPUTED)", xlab="Total Steps in a Day", col="red", border="black")

```

## Mean after imputing missing values
```{r}
 mean(totalperday_impute$steps)
```

## Median after imputing missing values
```{r}
median(totalperday_impute$steps)
```
Mean stays the same at 10766.19 but mean has changed slightly from 10765 to 10766.19

```{r}
#adding a day vector that specifies the day of the week
data$day<-weekdays(data$date)
#converting day to either weekday or weekend
for (i in 1:nrow(data)){
  if (data$day[i] %in% c("Monday", "Tuesday","Wednesday", "Thursday", "Friday"))
    data$day[i]<-"Weekday"
    
  else (data$day[i]<-"Weekend")
}
```


## Panel Plot of 5 minute interval and average number of steps taken across all weekdays and weekends

```{r}
#calculate mean of steps over weekday and weekend
avg_steps_weekday<-aggregate(steps~interval+day, data, FUN=mean)

library(ggplot2)
qplot(interval, steps, data=avg_steps_weekday,geom=c("line"),ylab="Number of Steps", xlab="5 Minute Interval")+facet_wrap(~day, ncol=1)
```